Overlay databank unlocks data-driven analyses of biomolecules for all

### 2024-01-12 main.tex
Editorial changes

Fig.~\ref{fig:quality} demonstrates how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for membrane simulations. Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08}, diverge between different force fields. To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Supplementary Figure~\ref{fig:top50simulations}), then picked force fields that occur in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C). Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) were discarded in this analysis. Because the average \textit{sn}-1 chain order parameter is a good proxy for the membrane packing (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A. Based on this, Lipid17 and Slipids simulations are most realistic for a POPC membrane, while CHARMM36 and GROMOS-CKP simulations predict overly packed bilayers (overestimated order in Supplementary Figure~\ref{fig:POPC_POPE_dataSI}). For POPE, on the other hand, GROMOS-CKP and Slipids are most realistic, while CHARMM36 and Lipid17 predict too packed membranes. In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant. Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}E--F and Supplementary Figure~\ref{fig:POPC_POPE_dataSI} a. Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation, POPC bilayer with OPLS3e parameters, for the reference.

@@

To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (areas per lipid) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~\ref{tab:MLpredictions}. The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes. Difference of 0.8\,nm in thickness is predicted between the thinnest (bacterial) and thickest (plasma) membranes. Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids Databank can be used to give valuable insights on biologically relevant properties of complex biological membranes. Most importantly, the delivered programmatic access to increasing amount of MD simulation data enables training of machine learning models that predict various membrane properties for all.

    \begin{table}[htb]
    \caption{Areas per lipid (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids Databank for membrane compositions corresponding different biological membranes. Systems are sorted in the order of increasing membrane packing. Compositions of different membranes are estimated based on references given in the composition column.}
        \centering
        \begin{tabular}{cccc}
             Membrane&  Composition [lipid(\%-fraction)]&  APL[\AA$^2$] & Thickness [nm$^2$]\\
             \hline
             Mitochondria & POPC(37) : POPE(31) : PI(6) : CL(22) : 
             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.4 \\
             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 61 (60)$^*$ & 4.2 \\
             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 57 & 4.6 \\
             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 52 & 4.8 \\
             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 5.0 \\
             Synaptic & POPC(28) : POPE(20) : POPS(9) : PI(2) : SM(4) : CHOL(37)~\cite{binotti21}  &  45 & 4.8 \\
             Influenza & POPC(5) : POPE(32) : POPS(15) : SM(9) : CHOL(40)~\cite{gerl12,ivanova15} & 42 & 4.8 \\
        \end{tabular}
        \label{tab:MLpredictions}

@@

To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids Databank using the general workflow depicted in Fig.~\ref{fig:overlay}D. The resulting non-zero values range between 0.3--322\,\textmu{}m/s with the mean of 14\,\textmu{}m/s and median of 8\,\textmu{}m/s. These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}. Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order. As expected, the permeability increases with the temperature, giving an average energy barrier of 14\,$\pm$\,3\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B. On the other hand, the water permeability on average decreases when membranes become more packed, that is, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability}C--E). Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability. Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters. Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Fig.~\ref{fig:permeationSI}).
%
%%Several models on how water dynamics depends on membrane properties have been proposed~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.

To examine how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we analyzed the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflows depicted in Figs.~\ref{fig:overlay}D and E. The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and~\ref{fig:diffusionSI}. Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately 50\% higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}. To estimate the diffusion anisotropy of water, $D_\perp/D_\parallel$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\perp$, using the Tanner equation~\cite{tanner78,wasterby02}. The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than the lateral diffusion coefficients of water (Figs.~\ref{fig:permeability}G--H), which is at the upper limit of anisotropy estimated from experimental data~\cite{nitsche19}. A significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ deviates further from unity with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability}G--H. This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability}C--D), while lateral diffusion remains approximately constant (Figs.~\ref{fig:diffusionSI}A,~C).

@@

\begin{table}[t]
    \centering
    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
     Target group  &  Outcomes     & Practical examples \\
    \hline
    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~\ref{tab:MLpredictions}), optimizing force field and model parameters from atomistic to continuum scale \\

@@

\subsection*{Experimental data}
Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids Databank repository. Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment. The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Table~\ref{tab:READMEkeysEXP} and in the NMRlipids Databank documentation~\cite{NMRlipidsExperimentalData}. The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}. In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in  Supplementary Figures~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.

\subsection*{Analysing simulations}
In practise, simulations in the NMRlipids Databank can be analyzed by executing a program that
1) loops over the README.yaml files in the {\it Databank layer},
2) downloads the data using the information in the README.yaml files, and then
3) performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in the README.yaml and mapping files.
This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a templates for user-defined analyses are available via NMRlipids Databank documentation~\cite{NMRlipidsExampleAndTutorials}. Further practical examples of codes performing such analyses are listed in Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.  The equilibration period given by the user (TIMELEFTOUT in Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes. For further details, see the NMRlipids Databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.

@@

Also, a quality measure based on differences in the simulated and experimental form factors across the available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}. However, the lobe heights in the simulated form factors depend on the simulation box size, as shown in Fig.~\ref{fig:sizedependence}; consequently, the quality measure defined in SIMtoEXP  would also depend on the simulation box size. In contrast, locations of the form factor minima (or, in precise terms: the minima of the absolute value of the form factor) are independent of the simulation box size (Fig.~\ref{fig:sizedependence}).
Here we use only the location of the first form factor minimum for quality evaluation,
because (due to fluctuations) the location of the second minimum is difficult to detect automatically in some experimental data sets, such as the POPE data in Figs.~\ref{fig:quality}D, E.
The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality}F), although the correlation of the second minima would be even stronger (Fig.~\ref{fig:QualityCorrelationsSI}).
In practise, we first filter the fluctuations from the form factor data using the Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minimum at  $q>$\,0.1\,\AA$^{-1}$ from both simulation ($F\!F_\mathrm{min}^\mathrm{sim}$) and experiment ($F\!F_\mathrm{min}^\mathrm{exp}$).
The quality of a form factor is then defined as the Euclidean distance between the minima locations: $F\!F_q = |F\!F_\mathrm{min}^\mathrm{sim}-F\!F_\mathrm{min}^\mathrm{exp}| \times 100$.

@@

  \noindent\begin{minipage}{\textwidth}
  \textbf{Table 2.} Examples of impact of NMRlipids and other overlay databanks in different disciplines.
    \centering
    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
     Target group  &  Outcomes     & Practical examples \\
    \hline
    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~\ref{tab:MLpredictions}), optimizing force field and model parameters from atomistic to continuum scale \\

>>>>

Fig.~\ref{fig:quality} demonstrates how the automatic simulation-quality evaluation and the NMRlipids Databank-API enable rapid selection of the best models for membrane simulations. Fig.~\ref{fig:quality}A illustrates that predictions for the lateral packing of membranes composed of two most biologically-abundant neutral membrane lipids, POPC and POPE~\cite{vanmeer08}, diverge between different force fields. To find the most realistic parameters to simulate membranes with these lipids, we first ranked all simulations based on order parameter quality (Supplementary Figure~\ref{fig:top50simulations}), then picked force fields that occur in Fig.~\ref{fig:quality}A (that is: force fields for both POPC and POPE in the Databank), and then ranked them according to the quality of the \textit{sn}-1 chain of POPC (Fig.~\ref{fig:quality}B) and of POPE (Fig.~\ref{fig:quality}C). Simulations with $\tau_\mathrm{rel}$ clearly above one (larger than 1.3) were discarded in this analysis. Because the average \textit{sn}-1 chain order parameter is a good proxy for the membrane packing (Fig.~\ref{fig:quality}G), rankings in Figs.~\ref{fig:quality}B and C can be used to select the simulations giving the most realistic results in Fig.~\ref{fig:quality}A. Based on this, Lipid17 and Slipids simulations are most realistic for a POPC membrane, while CHARMM36 and GROMOS-CKP simulations predict overly packed bilayers (overestimated order in Supplementary Figure~\ref{fig:POPC_POPE_dataSI}). For POPE, on the other hand, GROMOS-CKP and Slipids are most realistic, while CHARMM36 and Lipid17 predict too packed membranes. In conclusion, the quality evaluation based on the NMRlipids Databank suggests that the Slipids parameters are the best currently available choice for simulations with PC and PE lipids, at least for applications where membrane packing is relevant. Also direct comparisons with the experimental data for the most relevant simulations are shown in Figs.~\ref{fig:quality}D--F and Supplementary Figure~\ref{fig:POPC_POPE_dataSI} a. Fig.~\ref{fig:quality}D shows the overall highest-ranked simulation, POPC bilayer with OPLS3e parameters, for the reference.

@@

To demonstrate the usefulness of the constructed models for understanding multi-component membrane properties, we predicted later packing (areas per lipid) and membrane thicknesses of common biological membranes based on their lipid compositions reported in the literature, see table~1. The model predicts substantial 50\% difference in area per lipid between most densely (influenza virus) and loosely (mitochondria) packed membranes. Difference of 0.8\,nm in thickness is predicted between the thinnest (bacterial) and thickest (plasma) membranes. Such differences are expected to effect on many biologically relevant functions of membranes, such as permeation, cholesterol flip-flops (see next sections), and interactions with proteins~\cite{phillips09}, demonstrating that NMRlipids Databank can be used to give valuable insights on biologically relevant properties of complex biological membranes. Most importantly, the delivered programmatic access to increasing amount of MD simulation data enables training of machine learning models that predict various membrane properties for all.

    \begin{table}[htb]
    \caption{Areas per lipid (APL) and membrane thicknesses predicted by the linear regression model trained using the NMRlipids Databank for membrane compositions corresponding different biological membranes. Systems are sorted in the order of increasing membrane packing. Compositions of different membranes are estimated based on references given in the composition column.}\label{tab:MLpredictions}
        \centering
        \begin{tabular}{cccc}
             Membrane&  Composition [lipid(\%-fraction)]&  APL[\AA$^2$] & Thickness [nm$^2$]\\
             \hline
             Mitochondria & POPC(37) : POPE(31) : PI(6) : CL(22) : 
             CHOL(4)~\cite{vanmeer08,escriba15,casares19} & 74 (66)$^*$ & 4.4 \\
             Bacterial & POPC(20) : POPE(35) : POPG(35) : CL(5) : CHOL(5)~\cite{chwastek20}  & 61 (60)$^*$ & 4.2 \\
             ER &  POPC(54) : POPE(20) : POPS(4) : PI(11) : SM(4) : CHOL(8)~\cite{vanmeer08,escriba15,casares19} & 57 & 4.6 \\
             Golgi & POPC(36) : POPE(21) : POPS(6) : PI(12) : SM(7) : CHOL(18)~\cite{vanmeer08,escriba15,casares19} & 52 & 4.8 \\
             Plasma & POPC(23) : POPE(11) : POPS(8) : PI(7) : SM(17) : CHOL(34)~\cite{vanmeer08,escriba15,casares19} & 45 & 5.0 \\
             Synaptic & POPC(28) : POPE(20) : POPS(9) : PI(2) : SM(4) : CHOL(37)~\cite{binotti21}  &  45 & 4.8 \\
             Influenza & POPC(5) : POPE(32) : POPS(15) : SM(9) : CHOL(40)~\cite{gerl12,ivanova15} & 42 & 4.8 \\
        \end{tabular}

@@

To this end, we first calculated the water permeability through membranes from all simulations in the NMRlipids Databank using the general workflow depicted in Fig.~\ref{fig:overlay}D. The resulting non-zero values range between 0.3--322\,\textmu{}m/s with the mean of 14\,\textmu{}m/s and median of 8\,\textmu{}m/s. These values agree with the previously reported simulation results~\cite{venable19,camilo2022}, but are on average larger than experimental values reported for PC lipids in the liquid crystalline phase,~0.19--0.33\,\textmu{}m/s~\cite{jansen95}. Using the workflow depicted in Fig.~\ref{fig:overlay}E, we then plotted the observed permeabilities and their histogrammed values in Figs.~\ref{fig:permeability}B--E as a function of temperature, membrane thickness, area per lipid, and acyl chain order. As expected, the permeability increases with the temperature, giving an average energy barrier of 14\,$\pm$\,3\,$k_\mathrm{B}T$ for the water permeation from the Arrhenius plot in Fig.~\ref{fig:permeability}B. On the other hand, the water permeability on average decreases when membranes become more packed, that is, with decreasing area per lipid and increasing thickness and acyl chain order (Figs.~\ref{fig:permeability}C--E). Permeation of water through bilayers depends on membrane properties also according to previous studies, but there is no established consensus on whether the area per lipid~\cite{nagle08} or bilayer thickness~\cite{frallicciardi22} is the main parameter determining the permeability. Our analysis over the NMRlipids Databank, containing significantly more data than what was available in previous studies, suggest non-linear dependencies on both of these parameters. Clear dependencies of permeability on hydration level or the fraction of charged lipids, cholesterol, or POPE in the membrane were not observed (Supplementary Figure~\ref{fig:permeationSI}).
%
%%Several models on how water dynamics depends on membrane properties have been proposed~\cite{mathai07,nitsche13,nitsche16,shinoda16,venable19,frallicciardi22}.

To examine how water diffusion anisotropy depends on membrane properties in a multi-lamellar lipid bilayer system, we analyzed the water diffusion parallel to the membrane surface from all simulations in the NMRlipids Databank using the general workflows depicted in Figs.~\ref{fig:overlay}D and~\ref{fig:resultsProcessingSI}. The parallel diffusion coefficient of water, $D_\parallel$, decreases with reduced hydration and increases with the temperature, but dependencies on the membrane area per lipid, thickness, or fraction of charged lipids were not observed in Figs.~\ref{fig:permeability} and Supplementary Figure~\ref{fig:diffusionSI}. Simulation results are close to the experimental values with low hydration levels in Fig.~\ref{fig:permeability}F, but increase to approximately 50\% higher than the experimental value for bulk water diffusion value (3.1$\times$\,10$^{-9}$\,m$^2$/s at 313\,K~\cite{khakimov08}) with high hydration levels.
This is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion~\cite{pathirannahalage21}. To estimate the diffusion anisotropy of water, $D_\perp/D_\parallel$, in multilamellar membrane system, the permeability coefficients of water through membranes were translated to perpendicular diffusion coefficients, $D_\perp$, using the Tanner equation~\cite{tanner78,wasterby02}. The resulting perpendicular diffusion coefficients are approximately five orders of magnitude smaller than the lateral diffusion coefficients of water (Figs.~\ref{fig:permeability}G--H), which is at the upper limit of anisotropy estimated from experimental data~\cite{nitsche19}. A significant increase in the diffusion anisotropy with membrane packing is observed, as $D_\perp/D_\parallel$ deviates further from unity with decreasing area per lipid and increasing thickness in Figs.~\ref{fig:permeability}G--H. This follows from decreasing water permeability with membrane packing (Figs.~\ref{fig:permeability}C--D), while lateral diffusion remains approximately constant (Supplementary Figures.~\ref{fig:diffusionSI}A,~C).

@@

\begin{table}[t]
    \centering
    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
     Target group  &  Outcomes     & Practical examples \\
    \hline
    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~1), optimizing force field and model parameters from atomistic to continuum scale \\

@@

\subsection*{Experimental data}
Experimental data used in the quality evaluation, currently composed of C--H bond order parameters and X-ray scattering form factors, are stored in \texttt{/Data/experiments} in the NMRlipids Databank repository. Similarly to simulations, each experimental data set has a README.yaml file containing all the relevant information about the experiment. The keys and their descriptions for the experimental data, as well as detailed instructions to add the date, are given in Supplementary Table~\ref{tab:READMEkeysEXP} and in the NMRlipids Databank documentation~\cite{NMRlipidsExperimentalData}. The NMR data currently in the NMRlipids Databank are taken from Refs.~\citenum{scherer87,ferreira13,antila19,melcr20,bacle21,hanashima19,doktorova20} and the X-ray scattering data from Refs.~\citenum{Kucerka08a,kucerka11,pan12b,pan14,kucerka15,NMRlipidsIII,doktorova20}. In addition, previously unpublished NMR data for POPE, POPG, and DOPC was acquired as described in  Supplementary Figures~\ref{POPEexp1}--\ref{fig:DOPCexp}) and contributed to the Databank.

\subsection*{Analysing simulations}
In practise, simulations in the NMRlipids Databank can be analyzed by executing a program that
1) loops over the README.yaml files in the {\it Databank layer},
2) downloads the data using the information in the README.yaml files, and then
3) performs the desired analysis on a local computer utilising the universal naming conventions for molecules and atoms defined in the README.yaml and mapping files.
This general procedure is illustrated in Fig.~\ref{fig:overlay}D, and a templates for user-defined analyses are available via NMRlipids Databank documentation~\cite{NMRlipidsExampleAndTutorials}. Further practical examples of codes performing such analyses are listed in Supplementary Tables~\ref{tab:codes} and~\ref{tab:codesApplication}.  The equilibration period given by the user (TIMELEFTOUT in Supplementary Table~\ref{tab:READMEkeys}) is discarded from the trajectories in all analysis codes. For further details, see the NMRlipids Databank documentation \href{https://nmrlipids.github.io/}{nmrlipids.github.io}.

@@

Also, a quality measure based on differences in the simulated and experimental form factors across the available $q$-range is defined in the SIMtoEXP program~\cite{kucerka10}. However, the lobe heights in the simulated form factors depend on the simulation box size, as shown in Supplementary Figure~\ref{fig:sizedependence}; consequently, the quality measure defined in SIMtoEXP  would also depend on the simulation box size. In contrast, locations of the form factor minima (or, in precise terms: the minima of the absolute value of the form factor) are independent of the simulation box size (Supplementary Figure~\ref{fig:sizedependence}).
Here we use only the location of the first form factor minimum for quality evaluation,
because (due to fluctuations) the location of the second minimum is difficult to detect automatically in some experimental data sets, such as the POPE data in Figs.~\ref{fig:quality}E, F.
The first minimum correlates well with the thickness of a membrane (Fig.~\ref{fig:quality}G), although the correlation of the second minima would be even stronger (Supplementary Figure~\ref{fig:QualityCorrelationsSI}).
In practise, we first filter the fluctuations from the form factor data using the Savitzky--Golay filter (window length 30 and polynomial order 1) and locate the first minimum at  $q>$\,0.1\,\AA$^{-1}$ from both simulation ($F\!F_\mathrm{min}^\mathrm{sim}$) and experiment ($F\!F_\mathrm{min}^\mathrm{exp}$).
The quality of a form factor is then defined as the Euclidean distance between the minima locations: $F\!F_q = |F\!F_\mathrm{min}^\mathrm{sim}-F\!F_\mathrm{min}^\mathrm{exp}| \times 100$.

@@

  \noindent\begin{minipage}{\textwidth}
  \textbf{Table 2.} Examples of impact of NMRlipids and other overlay databanks in different disciplines.
    \centering
    \begin{tabular}{p{3.0cm}  p{4.0cm}  p{7.0cm}}
     Target group  &  Outcomes     & Practical examples \\
    \hline
    Computational and data scientists & Databanks with programmatic access, training sets for machine learning applications, automatic optimization of simulation models & Programmatic access to all publicly available MD simulation trajectories, machine learning models predicting properties of complex biomolecular assemblies (Table~1), optimizing force field and model parameters from atomistic to continuum scale \\
